### 2018-04-28 4_impl.tex
Set up the new structure of the impl chapter. Rewrite the beginning.

@@

Current Chapter describes the architecture of the \porthos[2] framework.

@@

\section{Requirements}
\label{ch:impl:req}

The main reason for starting the work on re-implementing the \porthos{} tool was the need for extension the input language and optimisation of the tool so that it is able to process real C programs in perspective.
In the existing \porthos{} architecture, the high-level recursive instructions (statements of C) are processed together with low-level non-recursive events abstractions (see classes of package `\texttt{dartagnan.program}' of \porthos) as single AST structure.
This AST was implemented as a mutable data structure, which is being modified during the stage of SMT-encoding.
We consider this architecture as hardly manageable (since it is hard to guarantee preserving of the invariants during the program execution) and poorly extensible (since adding support for a new high-level control-flow instruction requires changing multiple components of the program, from parser to encoder).
%Moreover, we encounter serious problems when we need to add support for control-flow jumps (as \texttt{continue}, \texttt{break}, \texttt{goto} in C).

Therefore, while designing the architecture of \porthos[2], we decided to clearly separate the high-level intermediate code representation (implemented as a recursive AST structure) from low-level event-based representation (implemented as an event-flow graph).
Such a modular architecture allows to use multiple input language parsers and convert parsed syntax trees to our AST, thus having support for multiple languages (for instance, the original input language of \porthos{} tool, two variants of syntax of a litmus test used by \tool{herd}, an assembly language for any supported architecture).

The following list enumerates the target \textit{requirements} for the new tool:

\begin{enumerate}[nolistsep]
    \item \textit{stability and transparency}
        \begin{itemize}
            \item following the principles of simplicity and readability;
            \item usage of software design patterns if necessary;
            \item usage of immutable data structures for all data transfer objects~(DTO);
            \item high code coverage by unit and functional tests;
        \end{itemize}
    \item \textit{efficiency}
        \begin{itemize}[leftmargin=1em]
            \item keeping the trade-off between execution time and memory usage;
        \end{itemize}
    \item \textit{extensibility}
        \begin{itemize}[leftmargin=1em]
            \item clear modular architecture
        \end{itemize}
\end{enumerate}

\section{Program Components}
\label{ch:impl:comp}

@@

\begin{figure}%[!b]%[H]
  \centering
  \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{img/my/lucidchart.com/data-flow-vert-300dpi.png}
  \caption{Main components of \porthos[2]}
  \label{fig:program-components}
\end{figure}

%\section{The input language}
%\label{ch:impl:input}
\subsection{Parsing the input}
\label{ch:impl:input-parser}

@@

    - syntactically only the integers were supported. we:arrays???pointers??enums?structs?

    - no declarations (all shared variables are declared in the init section, local variables are not decalred ever)

@@

\section{The Y-tree: an AST}
\label{ch:impl:ytree}

@@

\section{Model parser}
\label{ch:impl:model}

@@

@@

\subsection{Compilation}
\label{ch:impl:y2x:compil}

@@

\subsection{Post-compilation transformations}
\label{ch:impl:x2y:postcomp}

@@

\section{XGraph to ZFormula (SMT) encoder}
\label{ch:impl:comp:zformula}

@@

\section{Optimisations}

... performed on each stage

>>>>

The main reason for commencing the work on \porthos[2] was the need for processing real-world C programs, which, at first, requires extension the input language.
%process more intricate entities of the C language. 
%Extension of the input language imply not only ability to process syntactic sugar
This implies support not only for new syntactic structures of C language (such as the \texttt{switch} statement or the postfix increment operator \texttt{i++}), but also for its fundamental concepts and features (such as pointer arithmetic or first-order functions), which requires revision of the whole architecture of the tool.
Although far not the entire C language is supported (that, considering its complexity and numerous pitfalls, goes far beyond current thesis%
\footnote{To ensure that, we have merely to look at existing C compilers, for instance, gcc, that uses the hand-written C parser in more than 18.5 thousand lines (see \url{https://github.com/gcc-mirror/gcc/blob/master/gcc/c/c-parser.c})}%
), we consider the accomplished work as a step towards this.
%This applies not only to instances that simply constitute the syntactic sugar of C language (such as the \texttt{switch} statement or the prefix increment operator \texttt{++i}), but also to its fundamental concepts and features (such as pointer arithmetic or structures).
%However, the architecture of current version of \porthos{} does not allow to ... as it is ....
% Although and optimisation of the tool so that it is able to process real C programs in perspective.

%In current Chapter we discuss the new architecture of \porthos[2].

@@

%\section{The architecture of \porthos[1]}
\section{General principles}
\label{ch:impl:principles}

The existing implementation of \porthos{} does not distinguish the event-based program model from the high-level AST, they both are encoded into single SMT-formula (see classes of package `\texttt{dartagnan.program}' of \porthos tool).
Moreover, the syntax tree was implemented as a mutable data structure, which is being modified at all stages of the program (for instance, see the methods `\texttt{dartagnan.program.Program.compile(...)}' of \porthos{} that recursively compute some properties of the AST and change its state).
We are inclined to consider this architecture as one that is fast to develop, but hard to maintain (since it is difficult to guarantee the correctness of the program) and extend (since adding the support for a new high-level instruction requires changing multiple components of the program, from parser to encoder).
%Moreover, we encounter serious problems when we need to add support for control-flow jumps (as \texttt{continue}, \texttt{break}, \texttt{goto} in C).

% todo: in old porthos the strategy soft: print info about error. new: exception on any detected violation of any invariant

Therefore, while working on the new design of \porthos[2], we decided to clearly separate the high-level intermediate code representation (implemented as a recursive AST structure) from low-level event-based representation (implemented as an event-flow graph).
Such a modular architecture allows to support multiple input languages by parsing them and converting parsed syntax trees to a simplified AST. which, along with all other data-transfer objects (DTO), must be immutable, so that it is possible to guarantee the correctness of the program by controlling its invariants.
The immutability in \porthos[2] is implemented via \texttt{final} fields that are assigned by the immutable-object values (either a primitive type, or another immutable object, or an immutable collection provided by the library Guava by Google%
\footnote{Project repository: \url{https://github.com/google/guava/}}%
).

During the development of \porthos[2] we mainly followed the \textit{KISS principle}, which can be exhaustively described in 17 Unix Rules of Eric Raymond~\cite{raymond2003art}.% which include the rule of modularity, the rule of clarity, transparency, extensibility, etc.
The following list summarises the main rules we followed during the development of \porthos[2]:

%\vspace{1em}
\begin{enumerate}[nolistsep]
  \item \textit{robustness:}
    \begin{enumerate}[label*=\arabic*.]
      \item usage of immutable data structures for all DTOs;
      \item interrupting the work on errors;
      \item modular architecture: each module can be tested independently;
      \item usage of software design patterns if necessary;
    \end{enumerate}
  \item \textit{transparency:}
    \begin{enumerate}[label*=\arabic*.]
      \item following the principles of simplicity and readability;
      \item clear and informative program output;
    \end{enumerate}
  \item \textit{efficiency:}
    \begin{enumerate}[label*=\arabic*.]%[leftmargin=1.5em]
      \item keeping the trade-off between execution time and memory usage;
    \end{enumerate}
  \item \textit{extensibility:}
    \begin{enumerate}[label*=\arabic*.]%[leftmargin=1em]
      \item clear modular architecture
    \end{enumerate}
\end{enumerate}

\section{Architecture} %Program Components
\label{ch:impl:arch}

@@

@@

\begin{figure}%[!b]%[H]
  \centering
  % TODO: uncomment after a while
  %\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{img/my/lucidchart.com/data-flow-vert-300dpi.png}
  \caption{Main components of \porthos[2]}
  \label{fig:program-components}
\end{figure}

\subsection{The program input}

%(for instance, the original input language of \porthos{} tool, two variants of syntax of a litmus test used by \tool{herd}, an assembly language for any supported architecture).

\subsection{The Model}

blabla

\subsubsection{Y-tree}

\subsubsection{X-graph}

\subsubsection{W-model}

\subsubsection{Z-formula}

\subsection{The processing units} %Transformation}

\subsubsection{Y-tree parser}

\subsubsection{W-model parser}

\subsubsection{X-graph compiler}

\paragraph{Pre-compiler}
\paragraph{Compiler}
\paragraph{Post-compiler}

\subsubsection{The Z-formula encoder}

\subsection{The program output}

%\subsection{Parsing the input}
%\label{ch:impl:input-parser}

@@

    - syntactically only the integers were supported. we:arrays???pointers??enums?structs?
    - no declarations (all shared variables are declared in the init section, local variables are not decalred ever)

@@

%\section{The Y-tree: an AST}
%\label{ch:impl:ytree}

@@

%\section{Model parser}
%\label{ch:impl:model}

@@

\begin{figure}%[!b]%[H]
  \centering
  \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{img/my/class-diagrams/XEntity.png}
  \caption{TODO}
  \label{fig:class-diagrams:XEntity}
\end{figure}

\lstinputlisting{inc/XEventVisitor-cleaned.java}

@@

%\subsection{Compilation}
%\label{ch:impl:y2x:compil}

@@

%\subsection{Post-compilation}
%\label{ch:impl:x2y:postcomp}

transformations

@@

%\section{XGraph to ZFormula (SMT) encoder}
%\label{ch:impl:comp:zformula}

@@

%TODO: \section{Optimisations}
